---
title: 'Individual #1'
author: "Jacob Shoemaker; jshoemaker2249"
date: "4/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```

are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?
what time of the day do accidents happen (use variable HOUR)?
what is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

```{r}
ggplot(acc, aes(x=DAY_WEEK))+ geom_histogram() + scale_x_continuous(breaks = seq(1,7,1), labels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")) + xlab("Day of the Week") + ylab("Number of Accidents")
```


More accidents happen on Saturday and Sunday than any other days of the week.

```{r}
library(dplyr)
dat <- filter(acc, acc$HOUR <= 24)
ggplot(dat, aes(x=HOUR)) + geom_histogram() + scale_x_continuous(breaks = seq(0,24,1)) + xlab("Hour of Day") + ylab("Number of Accidents")
```


The most accidents happen in the evening, between 6pm and 9pm.

```{r}
drunk <- filter(acc, acc$DRUNK_DR >= 1)
drunk %>% summarise(numAcc = n())
```

There have been 8,769 accidents with at least one drunk driver.