---
title: "team"
author: "Jacob Shoemaker; jshoemaker2249, Qinwen Yang; AngelQinwen"
date: "4/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(tidyverse) 
library(lubridate)
library(maps)

acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```

are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?
what time of the day do accidents happen (use variable HOUR)?
what is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

```{r}

Dayweek <- acc %>% group_by(DAY_WEEK) %>% summarise(n=n())
Dayweek$DAY_WEEK = as.factor(c("Sunday","Monday","Tuesday", "Wednesday", "Thursday", "Friday","Saturday"))
ggplot(Dayweek, aes(x=DAY_WEEK, y=n, fill=DAY_WEEK))+geom_bar(stat="identity", width = 0.5)+xlab('Day of Week')+ylab('Number of Accident')+ggtitle('Relationship between Day of Week and Number of Accident ')+ scale_fill_manual(values = c("red","grey","red","red","grey","grey","grey"))+theme(axis.text.x = element_text(angle=90, hjust=1))
```


From the barplot, we can conclude that Friday, Saturday and Sunday of a week have more accidents happen than others.

```{r}

dat <- filter(acc, acc$HOUR <= 24)
ggplot(dat, aes(x=HOUR)) + geom_histogram() + scale_x_continuous(breaks = seq(0,24,1)) + xlab("Hour of Day") + ylab("Number of Accidents")
```


The most accidents happen in the evening, between 6pm and 9pm.

```{r}
drunk <- filter(acc, acc$DRUNK_DR >= 1)
drunk %>% summarise(numAcc = n())
```

There have been 8,769 accidents with at least one drunk driver.


Load the person table. Identify drivers (PER_TYP == 1, see fars manual) and subset on them.
Join accident and driver table (work out which variable(s) to use)

```{r}
drivers <- subset(per, PER_TYP ==1)
dat <- join(drivers, acc, type= "inner")
```


Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!
```{r}
dat %>% group_by(DAY_WEEK) %>% summarize(count = n())
ggplot(dat, aes(x=DAY_WEEK))+ geom_histogram() + scale_x_continuous(breaks = seq(1,7,1), labels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")) + xlab("Day of the Week") + ylab("Number of Accidents")
```


```{r}
filter(dat, dat$HOUR <= 24) %>% group_by(HOUR) %>% summarise(count = n()) 
dat1 <- filter(dat, dat$HOUR <= 24)
ggplot(dat1, aes(x = HOUR)) + geom_histogram() + scale_x_continuous(breaks = seq(0,24,1)) + xlab("Hour of Day") + ylab("Number of Accidents")
```


```{r}
dat %>% group_by(SEX) %>% summarise(count = n())
dat$SEX <- ifelse(dat$SEX == 8, 3, dat$SEX)
dat$SEX <- ifelse(dat$SEX == 9, 4, dat$SEX)
ggplot(dat, aes(x=SEX)) + geom_histogram() + scale_x_continuous(breaks = seq(1,4,1), labels = c("Male", "Female","Not Reported", "Unknown")) + ylab("Number of Accidents")
```

```{r}
JoinTab <- dat %>% filter(HOUR <= 24)
JoinTab$SEX<-factor(JoinTab$SEX)
levels(JoinTab$SEX) <- c("Male", "Female", "Not Reported","Unknown")
JoinTab$DAY_WEEK <- factor(JoinTab$DAY_WEEK)
levels(JoinTab$DAY_WEEK) <- c("Sunday","Monday","Tuesday","Wednesday", "Thursday","Friday","Saturday")
GroupJoinTab <- JoinTab %>% group_by(SEX, DAY_WEEK, HOUR) %>% summarise(n=n())
ggplot(GroupJoinTab, aes(x=SEX,y=n, fill=DAY_WEEK))+geom_bar(stat = "identity")+facet_wrap(~HOUR)+theme(axis.text.x = element_text(angle=90, hjust=1, size=10))+theme(axis.text=element_text(size=6))+ylim(c(0,400))+ylab('Number of Accidents')
  
```


Now plot a choropleth map of accidents on a county level
Read the description for the STATE and COUNTY columns in the FARS manual.
The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication
Use readxl::read_xlsx to read in the GLCs
Visualize the number of accidents in each county using a choropleth map. To do this, you need to join the accident data, GLCs, and appropriate map data.
Can you find seasonal trends in accident occurrences in different states? Is summer or winter more dangerous? Why?


Use readxl::read_xlsx to read in the GLCs
```{r}
#help(read_xlsx)
#str(GLCs)
GLCs <-  readxl::read_xlsx("FRPP GLC United States.xlsx", skip = 1)

```


Connecting Data and Make Maps:
Now plot a choropleth map of accidents on a county level
```{r}
glcs <- readxl::read_excel("FRPP GLC United States.xlsx", skip = 1)
library(tidyverse) 
library(lubridate)
library(maps)
counties <- map_data('county')

glcs1<- glcs
acc1 <- acc 
counties1 <- counties

glcs1$STATE <- glcs1$`State Code`
glcs1$COUNTY <- glcs1$`County Code`

glcs1 <- glcs1 %>% select("STATE", "State Name", "COUNTY", "County Name")
acc1 <- acc1 %>% select("STATE", "COUNTY", "ST_CASE")

geo1 <- acc1 %>% join(glcs1)
geo1 <- geo1 %>% distinct()
geo1$region <- geo1$`State Name`
geo1$subregion <- geo1$`County Name`
geo1 <- geo1 %>% select(-c("STATE", "COUNTY", "State Name", "County Name"))
geo1$region <- tolower(geo1$region)
geo1$subregion <- tolower(geo1$subregion)
glcs_acc <- geo1 %>% group_by(region, subregion) %>% summarise(numAcc = n())
final <- counties1 %>% join(glcs_acc)

ggplot(final, aes(x=long, y=lat, fill=numAcc)) + geom_polygon(aes(group=group)) + xlab("Longitude") + ylab("Latitude") + labs(fill = "Number of Accidents")
```

Can you find seasonal trends in accident occurrences in different states? Is summer or winter more dangerous? Why?
```{r}
acc <- acc %>% filter(WEATHER < 98)

Weather <- acc %>% group_by(STATE, WEATHER) %>% summarise(n=n())

ggplot(Weather, aes(x=STATE, y=n, fill=WEATHER))+geom_bar(stat = "identity")+xlim(c(0,60))+
  ylim(c(0, 3500))+ylab('Number of Accidents')

```


This bar char shows that the darker months (january and February most likely) account for a larger chunk of yearly accidents. This would make sense, as the winter weather can cause accidents. In addition, the lightly colored months (December and November) would also be subject to winter weather conditions.


